Are Delayed Issues Harder to Resolve?

### 2016-06-17 tex/responses.tex
Updates from ShareLaTeX

\begin{quote} {\em Thank you for the careful reviews. Using
those comments we have extensively revised and extended this draft.
We trust you will find this draft gratefully improved}
\end{quote}

@@

\begin{quote}{\em This new draft  now maps our work into the SE literature on theory building.}\end{quote}

@@

\begin{quote}{\em
This new draft  now maps not contains extensive notes on the origin and characteristics of our data.}\end{quote}

@@

\begin{quote}{\em All our results are now augmented with statistical significance tests as well as effect size tests. Those tests greatly strength the overall message of this paper-- that there is no significant
phase delay effect.}\end{quote}

@@

\begin{quote}{\em We have carefully read those comments  nd have made numerous changes
as directed by reviewer3.}\end{quote}

@@

\begin{quote}{\em  We have added an example of the DIE in the Introduction, and more precisely defined how we measure this effect Section 6.\footnote{Lucas- where exactly?} }\end{quote}

 The authors do not relate their reassessment of an
"old truism" to the growing set of papers on theory building
in software engineering. With access to such large amounts
of empirical evidence, it should be possible to take some
steps toward improved DIE theories.

\begin{quote}{\em  fshull XXX }\end{quote}

@@

Firstly, this draft now contains extensive notes on the different contexts of the
projects studied here. Those notes are offered in \tion{data_character} and \tion{171projects}.

Secondly (and this was the ineffective step), we  searched the numerous   listed in \tion{data_character} and \tion{171projects}
in order to check if any stratification contained the delayed issue effect. We failed to find a stratification
containing an exponential expansion in the time to fix issues. The reason for this was due to the dataL those
stratifications resulted in smaller sample sizes that exacerbated the large IQR's seen in this data
(to see these large IQRs,  please consider the 50th and IQR columns of Figure~\ref{fig:fix}.IQR is the
range 75th to 25th in the data. Note that most of our IQRs were larger than tan the 50th percentile; i.e. software
data exhibits large variances-- where are exacerbated by the smaller samples seen in the stratifications
 }\end{quote}

DETAILED REVIEW: The authors should expand on what they
include in the "delayed issue effect" (DIE), in particular
in relation to standard milestones in the life of an issue
and the corresponding timespans.  I'm particularly
interested in whether all "indirect" issue resolution steps
are covered, e.g., change impact analysis, updating
documentation, and regression testing. The effort involved
in these steps clearly depends on the development context of
the specific project. The short a-c) list on page 14:50
suggest that the issues studied are restricted to minor
fixes, i.e., no updates to architecture, changed hardware,
recertification, updated user manuals etc.

\begin{quote} {\em

We understand your comment here to be ``are we just reporting trivial errors?''
As shown in \fig{dtypes}, it is true that around a quarter of the fixes were simple documentation changes. That said, 75\% of the changes are quote  elaborate;   e.g. fixes to function necessitates a careful reflection of the purpose of the code.
}\end{quote}

@@

\begin{quote} {\em
At your suggestion, we had added extensive notes on the nature of our data--
see \tion{data_character} and \tion{171projects}.
}\end{quote}

@@

\begin{quote}{{\em 
Yes, there is undoubtedly some accuracy in the time-tracking. We provide some evidence on the TSP data set's integrity in Section 6.3, and a more extensive discussion on the issue of time-tracking and defect logging accuracy in Section 7.1 on Conclusion Validity.XXX lucas: please check the section numbers are currently
correct.
}
\end{quote}

The discussion on reassessing old truisms (Section 3) is
interesting, but it should be complemented by the
perspective of theory building in software engineering - An
active research topic lately, with a dedicated workshop
series (GTSE). I suggest looking into the following
references for a start: Sjøberg et al. (2008) "Building
Theories in Software Engineering", Smolander and Päivärinta
(2013) "Theorizing about software development practices",
and Stol and Fitzgerald (2015) "Theory-oriented software
engineering". Considering new empirical evidence is
obviously critical to theory building, and discussing your
new results in the light of an theory creation would be
valuable. The authors mention that DIE appears to occur
intermittently in certain kinds of projects - maybe the
authors could elaborate on this idea and present an improved
DIE theory based on what they now know? I believe the
authors have the best available data to do so, and I would
expect the paper to go beyond simply questioning the "old
truth".

 \begin{quote}{\em 
XXX Fshull
}\end{quote}

@@

\begin{quote}{\em 
 We have reviewed and cited the mentioned papers. We have incorporated many of the concerns raised in these papers into our expanded Section 7.4 on External Validity.
XXX lucas: please check the section numbers are currently
correct. XXX section 7.4 is a little... terse. can you add another 1/4 to 1/2 page of stuff? or is this sectionn just as large as it neeeds to be
}\end{quote}

The first and second claims are studied with much less
rigor. "DIE is a commonly held belief" is studied using a
survey of software engineers, both practitioners and
researchers. According to Fig. 2 (actually a table), the
number of respondents is 16 and 30 for practitioners and
researchers, respectively. Sixteen respondents from industry
represent a tiny survey of a very general claim that any
software engineer could respond to. Why were not more
answers collected? The authors do not have much evidence to
defend the first claim, and there is no discussion of the
corresponding validity threats. The second claim, "DIE is
poorly documented", is studied using a literature review.
Unfortunately, the method underlying the literature review
is not presented. Although it doesn't need to be an SLR of
the most rigorous kind, the authors should report how the
papers were identified. I suspect the terminology used to
describe the DIE phenomenon is highly diverse, thus it would
strengthen the paper if the authors reported how they
reviewed the literature. According to Page 19:34 only eight
publications were identified.

\begin{quote}{\em XXX
Lucas: need to show that these were the only empirical
results for DIE in tow
}\end{quote}

The iterative fashion of modern software development,
with agile at the extreme end, is not fully discussed (the
short discussion section could be extended). The phases of
the linear development of the 80s, (such as in Fig. 1)
probably don't exist in many of the projects in the TSP
dataset, still the authors discuss the DIE-effect from the
perspective of the 80s. Page 16:31 states "DIE has been used
to justify many changes to early lifecycle software
engineering" - does this mean the agile movement
successfully mitigated DIE? This possibility is not fully
considered in the paper.
How many of the 171 projects were
(more or less) agile? This appears to be an important
characteristic of the included projects - very important to
describe!

\begin{quote}{\em An astute observation. We provide more information on the types of projects included in our sample in Section 6.4. XXX Lucas. Check section number.
XXX Lucas: can't see the "percent agile" in 6.4. Can you work this with Bill?

More to the point -- does agile help mitigate DIE? The answer is - we don't know. A strong argument can be made that all iterative&incremental methodologies are meant to mitigate the DIE. Further, huge advances in software engineering technology and process have occurred since DIE was first observed, many of which have precisely targeted the DIE (static analysis, test-first development, automated build). Perhaps our analysis provides evidence that DIE can be defeated, at a large scale across multiple varied projects, using modern technologies. We do not yet know the causes here, but certainly it appears the DIE should not be treated as a truism any more, but rather a variable that can be controlled. We discuss this in Section 8.
}\end{quote}

@@

\begin{quote}{\em Those notes are offered in \tion{data_character} and \tion{171projects}. \end{quote}

Fig 1: "Operation" dwarfs everything in Boehm's diagram.
Since you do not study anything post release in this paper,
I think this figure skews the reader's mental picture. If
you remove the rightmost bar and rescale accordingly, the
plot better matches the findings you report in Fig. 10. You
still identify an interesting result though, but the
presentation turns fairer.

@@

\begin{quote}{\em We recently published a text on classic SE papers. We know how to manage those issues.}\end{quote}

@@

\begin{quote}{\em  We agree: that reference has been deleted. }\end{quote}

@@

\begin{quote}{\em  Our policy is to get one publication out of paper
before issueing the data in public. If this paper is accepted, we will certainly add that link. Note that we have a long history of living up to that policy-- see http://openscience.us/repo. }\end{quote}

@@

\begin{quote}{\em XXX Lucas. Not sure. but do you hanle all the following? }\end{quote}

@@

TYPOS ETC.  
Sec 1, §1: [3] and [30] have been swapped?  
Page 14:46 - Spell out IV\&V the first (and only) time it appears.
Page 3:24 - missing verb (is) Page 16:43 - "Did this study failed" 
Page 18:44 - Appache 
Page 19:19 - Al Ref 26 - oo --> OO

@@

\begin{quote}{\em 
Those notes are offered in \tion{data_character} and \tion{171projects}.}\end{qupte}

I think in particular about the size (lines of codes and/or
 number of software components); a duration histogram would
 also be relevant. Accordingly, the formula for calculating
 variable "Total effort" in Fig. 7 could be explained, and
 Figure 9 made more readable (bigger size).

 \begin{quote}{\em We have added more description of the projects to Section 6.4. We have also added more precise definitions of our measures in Section 6.2. We have adjusted the size of figures throughout the paper to be more readable.XXX Lucas please check section references. We have added more text 
 to explain {\em effort} in \fig{dist}.}\end{quote}

@@

\begin{quote}{\em 
Those notes are offered in \tion{data_character} and \tion{171projects}.}\end{qupte}

Figure 10 is central to the paper's demonstration; its
expressiveness could be enhanced: the formula for calculating the 50th
and 95 ratio could be provided,  the unit of column
"Percentile" could be mentioned.

\begin{quote}{\em
\fig{raw}'s caption now has extensive explanation text. e.g. units
are added on line two. Also, an example of the calculations of the ratios is offered
in the last line of the caption. }\end{quote}

@@

\begin{quote}{\em
Not explicitly, but \fig{dtypes} lists "interfaces" as a the fourth most
common type of defect. Note that "interfaces" are changed by architectural
issues. }\end{quote}

@@

\begin{quote}{\em
  Absolutely - there are potentially many explanations for why the DIE was NOT observed in our dataset. We discuss some more of the potential causes in Section 8.
  XXX Lucas. Is the section reference right? Do we reference Harter (and if we do, mention that here).?
 }\end{quote}

@@

\begin{quote}{\em
 XXX Lucas: can you retire the follwoing
 }\end{quote}

Minor remarks: - p.9, line 29 : " … reported by Shull [52], found that the cost to find certain non-critical classes of defects …" => is it the "cost to find" or the "cost to FIX"?  
```

- [X]  llayman It should be "cost to fix". We have corrected this.

- p.10, line 21: duplicate word "there" -
p.12, line19: missing word "One of THE guiding …" - p.16,
line 33: "distinguish" instead of "distinguished"?  - p.18,
line 42: explain acronym MEAN - p.20, line 20: useless
enumeration A1?

{\bf  Reviewer \#3's Comemnts}:

@@

The empirical analysis (Section 6) is a bit of a let down. I
would like to see some sort of more robust and rigorous
statistical analysis, based on statistical tests.

@@

Section 2

I'm not really sure what the authors mean by "We say that a
measure collected in phase 1, ., i, .. j is very much more
when that measure at phase j is larger than the sum of that
measure in earlier phases 1 <= i < j." You are defining a
property of a measure, but, the way this definition is
written, it seems as if you are defining the property of a
measure of "being much more," without any further
qualifications.  I can guess that you mean that a measure
$m$ (collected in phases, so you can denote by $m(i)$ the
value of $m$ in phase $i$) has this property *in phase $j$*
(so, it's a property m has in a specific phase and not a
general property of the measure) if $\sum_{1 <= i < i} m(i)
< m(j)$. Then, it's up to you to make this a property of the
measure, for example by using an existential "policy" ($m$
has this property if there exists one phase $j$ where
$\sum_{1 <= i < i} m(i) < m(j)$) or a universal one ($m$ has
this property if for all phases $1<j$, $\sum_{1 <= i < i}
m(i) < m(j)$).
\begin{quote}{\em
Thanks for that. We added the $m_j$ notation to Section 2, as you suggested.
 }\end{quote}

Your definition of "difficult issue" is "Issues are more
difficult when their resolution takes more time or costs
more (e.g. needs expensive debugging tools or the skills of
expensive developers)." That's not a very precise
definition, though, because it has two different
interpretations, one in terms of time and the other in terms
of effort. The reasons why you introduce this definition
become clearer only in Section 7.2, in the Threats to
Validity, and that's too late. You should move some of the
discussion of Section 7.2 here.
\begin{quote}
{\em Good idea! We did just that- see middle para in Section 2.}
\end{quote}
Why would the term "delayed issue effect" be a
generalization of the rule "requirements errors are hardest
to fix"? "Delayed" seems to be quite specific as it appears
to refer to time, while "hardest to fix" may refer to other
variables, like effort or cost.
\begin{quote}
{\em Not sure if this a reviewer requirement for something to be fixed
prior to publication, or a clarification question. We used these
terms since when we ran the survey, the audience clearly made
that generalization. Does this reply adequately to your comment?}
\end{quote}
In Claim 3, "very much more harder" sounds like a
little bit too much ... ("more harder"?)
\begin{quote}
{\em Our bad. We fixed that.}
\end{quote}
Section 4

A very minor issue: why is Figure 2 a ... figure, instead of
a table? Same for all other tables ...
\begin{quote}
{\em Authors' preference. When a document has many tables and figures, navigating
between them is hard. Calling them all "figures" makes navigation faster. 
But do you want us to change that?}
\end{quote}
Section 5.1

@@

You should rephrase your definition "a defect is any change," since a defect is what existed before the change was made and not the change itself.
\begin{quote}
{\em We now provide a more precise definition of defect in Section 6.2.
XXXX= Lucas: check section number.
}
\end{quote} 
You do not explain that "QualTest" in Fig. 8 is.
\begin{quote}
{\em Removed}
\end{quote}
The definition of "time per defect - The total # of defects found in a plan item during a removal phase divided by the total time spent on that plan item in that phase." is not correct as it is, as this would basically be the number of defects per unit time, instead. The roles of time and
defects should be reversed.
\begin{quote}{\em We now provide a more precise definition of time measurement related to defects in Section 6.2.}\end{quote}

Section 6.4

@@

Figure 9. The caption says "Distribution of defects found
and fixed by phase." Is that the distribution of defects
found in some phase and fixed in that phase? The text says
"The distribution of defects found and fixed per phase in
our data is shown in Figure 9. A high percentage of defects
(44\) were found and fixed in the early phases, i.e.,
requirements, high level design, and design reviews and
inspections." which doesn't add much. If that's the
distribution of defects introduced in a phase and fixed
immediately in that phase, it is not clear why you introduce
it, though.
\begin{quote}
{\em XXXX Bill: one for you.}
\end{quote}
Figure 10. The explanation of the data and histograms in the
figure is, at best, confusing. The caption says "50th and
95th percentiles of issue resolution times" and the opening
sentence of Section 6.4 says "Figure 10 shows the 50th and
95th percentile of the time spent resolving issues ..."
However, these are not "times," because you write "expressed
as ratios of resolution time in the phase where they were
first injected" a couple of paragraphs below. This seems to
be the right interpretation, but right after that you write
"The BLACK and RED bars show the increases for the 50th
(median) and 95th percentile values, respectively," which
would be a third interpretation of the data in Figure 10.
\begin{quote}
{\em \fig{raw} now has an extensive help text in the caption.
We used your comments there where we noted the ratios used in the
bard charts are unitless while the left-hand-side column numbers are
times in minutes.
}
\end{quote}

@@

Section 7.1

Maybe you could analyze the projects developed in a
"traditional" way and check if there is some sort of DIE
there. For example, if you look at the projects developed
with the waterfall life cycle, maybe you will find larger
DIE than for the other life cycles, which might partially
justify the claims of previous researchers about the
existence of DIE. This would also show that newer kinds of
life cycles (like Agile ones) help solve DIE. You do provide
hints about this in Section 8 ("We also note that other
development practices have changed in ways that could
mitigate the delayed issued effect") and Section 9, but
that's clearly not enough.
\begin{quote}{\em
If your comment is along the lines of ``look for stratifications to check if
DIE exists in subsets'' then we offer the following failed study.

We looked into those stratifications and have  failed to find a stratification
containing an exponential expansion in the time to fix issues. The reason for this was due to the data in  those
stratifications resulted in smaller sample sizes that exacerbated the large IQR's seen in this data
(to see these large IQRs,  please consider the 50th and IQR columns of Figure~\ref{fig:fix}.IQR is the
range 75th to 25th in the data. Note that most of our IQRs were larger than tan the 50th percentile; i.e. software
data exhibits large variances-- where are exacerbated by the smaller samples seen in the stratifications.}
\end{quote}
Section 8

- [ ] Verify that the Section #'s in our responses are accurate prior to submission.

>>>>

\begin{quote} {\em Thank you for the careful reviews. Using
those comments we have extensively revised and extended this draft.
Specifically, the old version was 20 pages and this version is 27 pages (and the new
material all comes from the clarification and correction points offered by the reviewers).
We hope that  you will find this draft greatly improved.}
\end{quote}

@@

\begin{quote}{\em This new draft now maps our work into the SE literature on theory building. This new material appears in the introduction as well as being
peppered throughout the document (in numerous places).}\end{quote}

@@

\begin{quote}{\em
This new draft contains much more detail on the origin and characteristics of our data (see \tion{data-collection} and \tion{data_character}).}\end{quote}

@@

\begin{quote}{\em All our results are now augmented with statistical significance tests as well as effect size tests (see \tion{stats} and \tion{171projects}). Those tests strengthen the overall message of this paper-- that there is no significant
phase delay effect.}\end{quote}

@@

\begin{quote}{\em We have carefully read those comments and have made numerous changes
as directed by reviewer3.}\end{quote}

@@

\begin{quote}{\em  The introduction now includes an example of DIE  (see Fig 1).
Also there are now more precise  definitions on how we measure this effect in \tion{data-collection}}\end{quote}

 The authors do not relate their reassessment of an
``old truism'' to the growing set of papers on theory building
in software engineering. With access to such large amounts
of empirical evidence, it should be possible to take some
steps toward improved DIE theories.

\begin{quote}{\em We now map our work to prior publications on theory-building in software engineering. That mapping appears in numerous comments through out the text (using the references offered by this reviewer) with the bulk of the comments being top of p2 and the conclusion.

Notably, we now introduce the DIE as an example of such a theory (in the introduction), and explain that this is done on the basis of it providing a description of a Software development phenomenon accompanied by a rationale that explains why such behavior is expected. While we introduce in the conclusions several rationales that can explain why the effect called for by the theory was not observed in our dataset, we feel that future work is needed to collect data in support or contradiction of these explanations before a new theory can be formulated with confidence. }\end{quote}

@@

First, this draft now contains extensive notes on the different contexts of the
projects studied here. Those notes are offered in \tion{data_character} and \tion{171projects}.

Second (and this was the ineffective step), we  searched the categories   listed in \tion{data_character} and \tion{171projects}
in order to check if any stratification contained the delayed issue effect. 
That study   failed to find a stratification
containing an exponential expansion in the time to fix issues. The reason for this was  the small size
of those stratifications  exacerbated the large IQR's seen in this data\footnote{Recall that in a sorted list of numbers,
the inter-quartile range, or IQR, is difference between the 
  75th to 25th percentile value.}:
\bi
\item
Our 171 projects   stratifies into subsets of varying sizes. Sorting those sizes, we get a list
comprising of 17,12 then numerous much smaller stratifications.  Reasoning over such small samples
is problematic in the general case and, in the case of our data, it is ever more problematic due to
the large IQRS of the data:
\item
To see these large IQRs,  please consider the 50th and IQR columns of Figure~\ref{fig:raw} Note that most of our IQRs were larger than tan the 50th percentile; i.e. software
data exhibits large variances-- where are exacerbated by the smaller samples seen in the stratifications.
\ei
Our conclusion from exploring the stratification's is, given the currently available data, we cannot check
for a DIE effect in subsets of this data.

Note that the above notes are now added to the paper at the end of Section 6.6.
 }\end{quote}

DETAILED REVIEW: The authors should expand on what they
include in the ``delayed issue effect'' (DIE), in particular
in relation to standard milestones in the life of an issue
and the corresponding timespans.  I'm particularly
interested in whether all "indirect" issue resolution steps
are covered, e.g., change impact analysis, updating
documentation, and regression testing. The effort involved
in these steps clearly depends on the development context of
the specific project. The short a-c) list on page 14:50
suggest that the issues studied are restricted to minor
fixes, i.e., no updates to architecture, changed hardware,
recertification, updated user manuals etc.

\begin{quote} {\em
We now provide a more explicit definition of what constitutes a defect (our study of issues is limited to defects) and the ``defect lifecycle'' phases we measure in \tion{data-collection}. These definitions greatly expand on the now-deleted list on 14:50 in teh previous version. In TSP, ``a defect is any change to a product, after its construction, that is necessary to make the product correct''. This includes both minor issues (typography) and severe issues (architecture changes, requirements errors). Time on a defect includes a) time to investigate/analyze a defect once discovered, b) time to craft and implement a fix, and c) time to verify and close the defect.

Regarding the concern that we may only be reporting trivial defects, we have added \fig{dtypes}, which shows the distribution of defect types in our data. It is true that around a quarter of the fixes were documentation changes, but the other 75\% are function-altering behavior in the system. 
}\end{quote}

@@

\begin{quote} {\em
At your suggestion, we have added extensive notes on the nature of our data--
see \tion{data_character} and \tion{171projects}.
}\end{quote}

@@

\begin{quote}{\em 
Yes, there is undoubtedly some accuracy in the time-tracking. We provide some evidence on the rigor of the TSP data collection process in \tion{defin}, \tion{data_integrity}, and \tion{data_character}, and a more extensive discussion on the issue of time-tracking and defect logging accuracy in \tion{conclusion_validity} on Conclusion Validity. We note that the data being collected here was not collected
for the purposes of this paper. Rather, it was being collected and used by the
programmers to assist them in their day-to-day management of the project.
Given this, we have confidence that the logging information is not widely inaccurate.

Please also consider the sanity checks reported in a companion paper (referenced here \tion{data_integrity}). In summary, this data has been heavily vetted.
}
\end{quote}

The discussion on reassessing old truisms (Section 3) is
interesting, but it should be complemented by the
perspective of theory building in software engineering - An
active research topic lately, with a dedicated workshop
series (GTSE). I suggest looking into the following
references for a start: Sj{\o}berg et al. (2008) ``Building
Theories in Software Engineering'', Smolander and Paivarinta
(2013) ``Theorizing about software development practices'',
and Stol and Fitzgerald (2015) "Theory-oriented software
engineering". Considering new empirical evidence is
obviously critical to theory building, and discussing your
new results in the light of an theory creation would be
valuable. The authors mention that DIE appears to occur
intermittently in certain kinds of projects - maybe the
authors could elaborate on this idea and present an improved
DIE theory based on what they now know? I believe the
authors have the best available data to do so, and I would
expect the paper to go beyond simply questioning the "old
truth".

 \begin{quote}{\em 
We thank the reviewer for this important observation and have connected our work to the literature on theory-building. However, our data supports a number of possible explanations for the observed intermittent appearance of the DIE. Since the current data is not suitable for choosing among the various hypothesized explanations we feel that formulating a new theory build from this data
would be premature at this time. That said, we think the current paper would be useful
as a data point which, when combined with other sources, could be useful for some new future theory of SE.
}\end{quote}

@@

\begin{quote}{\em 
 We have reviewed and cited the mentioned papers. We have incorporated many of the concerns raised in these papers into the expanded \tion{external_validity} on External Validity.
 There, we note that Sj{\o}berg et al. list three classes of threats to External Validity (the smallest number) and we address all three of them. In truth, the only way to address External Validity threats is to have a huge sample, broadly representative of your domain, over an extended period of time-- e,g. using the 171 projects from 2006 to 2014 studied in this paper. The issues external validity in observational studies raised in \cite{madigan2014} and \cite{carlson2009} are more applicable to studies that compare a treatment and a control, whereas we are observing an effect that leads us to question the theory that DIE is a truism.
}\end{quote}

The first and second claims are studied with much less
rigor. "DIE is a commonly held belief" is studied using a
survey of software engineers, both practitioners and
researchers. According to Fig. 2 (actually a table) [NOTE: Fig. 2 is now \fig{survey_results}], the
number of respondents is 16 and 30 for practitioners and
researchers, respectively. Sixteen respondents from industry
represent a tiny survey of a very general claim that any
software engineer could respond to. Why were not more
answers collected? The authors do not have much evidence to
defend the first claim, and there is no discussion of the
corresponding validity threats. The second claim, "DIE is
poorly documented", is studied using a literature review.
Unfortunately, the method underlying the literature review
is not presented. Although it doesn't need to be an SLR of
the most rigorous kind, the authors should report how the
papers were identified. I suspect the terminology used to
describe the DIE phenomenon is highly diverse, thus it would
strengthen the paper if the authors reported how they
reviewed the literature. According to Page 19:34 [NOTE: the enumerated list in the Conclusion] only eight
publications were identified.

\begin{quote}{\em We agree with the reviewer - the phrasing of our claims and the supporting evidence were not congruent. Our intention is to argue from multiple angles that the evidence for DIE is sparse and, more importantly, that DIE persists as a popular belief. Our goal here is to motivate that we should re-investigate the DIE. We have toned down the language and combined our arguments on ``widely-believed'' and ``poorly-documented'' into a new \tion{belief} whose focus is to motivate our study.}\end{quote}

The iterative fashion of modern software development,
with agile at the extreme end, is not fully discussed (the
short discussion section could be extended). The phases of
the linear development of the 80s, (such as in Fig. 1) [NOTE: Fig. 1 remains \fig{b81}] 
probably don't exist in many of the projects in the TSP
dataset, still the authors discuss the DIE-effect from the
perspective of the 80s. Page 16:31 states "DIE has been used
to justify many changes to early lifecycle software
engineering" - does this mean the agile movement
successfully mitigated DIE? This possibility is not fully
considered in the paper.
How many of the 171 projects were
(more or less) agile? This appears to be an important
characteristic of the included projects - very important to
describe!

\begin{quote}{\em An astute observation. We provide more information on the types of projects included in our sample in \tion{data_character}

More to the point -- does agile help mitigate DIE? The answer is - we don't know. A strong argument can be made that all iterative \& incremental methodologies were designed to mitigate the DIE. Further, huge advances in software engineering technology and process have occurred since DIE was first observed, many of which have precisely targeted the DIE (static analysis, test-first development, automated build). Perhaps our analysis provides evidence that DIE can be defeated, at a large scale across multiple varied projects, using modern technologies. We do not yet know the causes here, but certainly it appears the DIE should not be treated as a truism any more, but rather a variable that can be controlled. We have added more discussion on this to \tion{discussion}.
}\end{quote}

@@

\begin{quote}{\em Much more context is now offered in \tion{data_character} and \tion{171projects}}. \end{quote}

Fig 1 [NOTE: Still \fig{b81}]: "Operation" dwarfs everything in Boehm's diagram.
Since you do not study anything post release in this paper,
I think this figure skews the reader's mental picture. If
you remove the rightmost bar and rescale accordingly, the
plot better matches the findings you report in Fig. 10 [NOTE: now \fig{raw}]. You
still identify an interesting result though, but the
presentation turns fairer.

@@

\begin{quote}{\em We have removed all copyrighted images and replaced them with original drawings based on information from the cited source.}\end{quote}

@@

\begin{quote}{\em  We agree: that statement has been deleted. }\end{quote}

@@

\begin{quote}{\em  Our policy is to get one publication out of paper
before issuing the data in public. If this paper is accepted, we will certainly add that link. Note that we have a long history of living up to that policy-- see http://openscience.us/repo. }\end{quote}

@@

@@

\begin{quote}{\em We have removed this statement.}\end{quote}

TYPOS ETC.

Sec 1, §1: [3] and [30] have been swapped?

Page 14:46 - Spell out IV\&V the first (and only) time it appears.

Page 3:24 - missing verb (is) Page 16:43 - "Did this study failed"

Page 18:44 - Appache

Page 19:19 - Al Ref 26 - oo --> OO

\begin{quote}{\em  All of these have been corrected. }\end{quote}

@@

\begin{quote}{\em 
More description about the context and sample have been added in \tion{data_character} and \tion{171projects}.}\end{quote}

I think in particular about the size (lines of codes and/or
 number of software components); a duration histogram would
 also be relevant. Accordingly, the formula for calculating
 variable "Total effort" in Fig. 7 [NOTE: now \fig{dist}] could be explained, and
 Figure 9 [NOTE: now \fig{fix-phase-dist}] made more readable (bigger size).

 \begin{quote}{\em We have added more description of the projects in \tion{data_character}. We have also added more precise definitions of our measures in \tion{data-collection}. We have adjusted the size of figures throughout the paper to be more readable. We have added more text 
 to explain {\em effort} in \fig{dist}.}\end{quote}

@@

\begin{quote}{\em 
Those notes are offered in \tion{data_character} and \tion{171projects}.}\end{quote}

Figure 10 [NOTE: now \fig{raw}] is central to the paper's demonstration; its
expressiveness could be enhanced: the formula for calculating the 50th
and 95 ratio could be provided,  the unit of column
"Percentile" could be mentioned.

\begin{quote}{\em
\fig{raw}'s caption (formerly Figure 10) now has extensive explanation text. e.g. units
are added on line two. Also, an example of the calculations of the ratios is offered
in the last line of the caption. Further,
to simplify comprehension of this diagram, it has been split into two
(and the more arcane discussion about the outliers, statistics gathered, and methods used is provided is also provided).}\end{quote}

@@

\begin{quote}{\em
Not explicitly, but \fig{dtypes} lists ``interfaces'' as a the fourth most
common type of defect. Note that ``interfaces'' are changed by architectural
issues. }\end{quote}

@@

\begin{quote}{\em
  Absolutely - there are potentially many explanations for why the DIE was NOT observed in our dataset. We add more discussion on the potential causes in \tion{discussion}, supported by the reference to Harter et al.
 }\end{quote}

@@

Minor remarks:

- p.9, line 29 : " … reported by Shull [52], found that the cost to find certain non-critical classes of defects ..." => is it the "cost to find" or the "cost to FIX"?

- p.10, line 21: duplicate word "there" -

p.12, line19: missing word "One of THE guiding …"

- p.16, line 33: "distinguish" instead of "distinguished"?

- p.18, line 42: explain acronym MEAN

- p.20, line 20: useless enumeration A1?

\begin{quote}{\em
 All of these have been corrected or removed.
 }\end{quote}

{\bf  Reviewer \#3's Comments}:

@@

The empirical analysis (Section 6 [NOTE: now \tion{analysis}) is a bit of a let down. I
would like to see some sort of more robust and rigorous
statistical analysis, based on statistical tests.

@@

Section 2 [NOTE: still \tion{claims}]

I'm not really sure what the authors mean by "We say that a
measure collected in phase 1, ., i, .. j is very much more
when that measure at phase j is larger than the sum of that
measure in earlier phases 1 <= i < j." You are defining a
property of a measure, but, the way this definition is
written, it seems as if you are defining the property of a
measure of "being much more," without any further
qualifications.  I can guess that you mean that a measure
$m$ (collected in phases, so you can denote by $m(i)$ the
value of $m$ in phase $i$) has this property *in phase $j$*
(so, it's a property m has in a specific phase and not a
general property of the measure) if $\sum_{1 <= i < i} m(i)
< m(j)$. 
\begin{quote}{\em
Thanks for that. We added the $m_j$ notation to \tion{claims}, as you suggested.
 }\end{quote}

Your definition of "difficult issue" is "Issues are more
difficult when their resolution takes more time or costs
more (e.g. needs expensive debugging tools or the skills of
expensive developers)." That's not a very precise
definition, though, because it has two different
interpretations, one in terms of time and the other in terms
of effort. The reasons why you introduce this definition
become clearer only in Section 7.2 [NOTE: now \tion{construct}], in the Threats to
Validity, and that's too late. You should move some of the
discussion of Section 7.2 here.
\begin{quote}
{\em Good idea! We did just that- see middle para in \tion{claims}.}
\end{quote}
Why would the term "delayed issue effect" be a
generalization of the rule "requirements errors are hardest
to fix"? "Delayed" seems to be quite specific as it appears
to refer to time, while "hardest to fix" may refer to other
variables, like effort or cost.
\begin{quote}
{\em Not sure if this a reviewer requirement for something to be fixed
prior to publication, or a clarification question. We used these
terms since when we ran the survey, the audience clearly made
that generalization. Does this reply adequately to your comment?}
\end{quote}
In Claim 3, "very much more harder" sounds like a
little bit too much ... ("more harder"?)
\begin{quote}
{\em Our bad. We fixed that.}
\end{quote}
Section 4

A very minor issue: why is Figure 2 [NOTE: now \fig{survey_results}] a ... figure, instead of
a table? Same for all other tables ...
\begin{quote}
{\em Authors' preference. When a document has many tables and figures, navigating
between them is hard. Calling them all "figures" makes navigation faster. 
But do you want us to change that?}
\end{quote}
Section 5.1

@@

You should rephrase your definition "a defect is any change," since a defect is what existed before the change was made and not the change itself.
\begin{quote}
{\em We now provide a more precise definition of defect in \tion{defin}.
}
\end{quote} 
You do not explain that "QualTest" in Fig. 8 [NOTE: now \fig{waterfall}] is.
\begin{quote}
{\em A mistake on our part that has been removed from the figure.}
\end{quote}
The definition of "time per defect - The total \# of defects found in a plan item during a removal phase divided by the total time spent on that plan item in that phase." is not correct as it is, as this would basically be the number of defects per unit time, instead. The roles of time and
defects should be reversed.
\begin{quote}{\em We now provide a more precise definition of time measurement related to defects in \tion{data-collection}}\end{quote}

@@

Figure 9 [NOTE: now \fig{fix-phase-dist}]. The caption says "Distribution of defects found
and fixed by phase." Is that the distribution of defects
found in some phase and fixed in that phase? The text says
"The distribution of defects found and fixed per phase in
our data is shown in Figure 9. A high percentage of defects
(44) were found and fixed in the early phases, i.e.,
requirements, high level design, and design reviews and
inspections." which doesn't add much. If that's the
distribution of defects introduced in a phase and fixed
immediately in that phase, it is not clear why you introduce
it, though.
\begin{quote}
{\em 
The TSP convention is to fix a defect in the phase in which it was identified, thus "find and fix". With some exceptions, defects are not counted as defects unless they escape that phase. The exception case is the secondary defect introduced during a fix that occurs with some frequency in test and rarely in inspection or review. We have clarified this distinction in the text.}
\end{quote}
Figure 10 [NOTE: now \fig{raw}]. The explanation of the data and histograms in the
figure is, at best, confusing. The caption says "50th and
95th percentiles of issue resolution times" and the opening
sentence of Section 6.4 says "Figure 10 shows the 50th and
95th percentile of the time spent resolving issues ..."
However, these are not "times," because you write "expressed
as ratios of resolution time in the phase where they were
first injected" a couple of paragraphs below. This seems to
be the right interpretation, but right after that you write
"The BLACK and RED bars show the increases for the 50th
(median) and 95th percentile values, respectively," which
would be a third interpretation of the data in Figure 10.
\begin{quote}
{\em We agree-that figure was far too confusing. We have now split the 50th and 90th peercentile results
into two charts, which simplifies both. Please see \fig{raw} and \fig{last}.
Also, please note that that fit\fig{raw} now has an extensive help text in the caption.
We used your comments there where we noted the ratios used in the
bard charts are unitless while the left-hand-side column numbers are
times in minutes.
}
\end{quote}

@@

Maybe you could analyze the projects developed in a
"traditional" way and check if there is some sort of DIE
there. For example, if you look at the projects developed
with the waterfall life cycle, maybe you will find larger
DIE than for the other life cycles, which might partially
justify the claims of previous researchers about the
existence of DIE. This would also show that newer kinds of
life cycles (like Agile ones) help solve DIE. You do provide
hints about this in Section 8 [NOTE: now \tion{discussion}] ("We also note that other
development practices have changed in ways that could
mitigate the delayed issued effect") and Section 9 [NOTE: now \tion{conclusion}], but
that's clearly not enough.
\begin{quote}{\em

If your comment is along the lines of ``look for stratifications to check if
DIE exists in subsets'' then we offer the following notes (which we added to the end of \tion{171projects}):
